\documentclass{article}

\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[margin=0.5in]{geometry}
\usepackage[hidelinks, bookmarks=false]{hyperref}

\usepackage[dvipsnames]{xcolor}

\let\oldemptyset\emptyset
\let\emptyset\varnothing
\newcommand{\R}{\mathbb{R}}
\newcommand{\Rd}{\R^d}

\usepackage{environ}
\NewEnviron{centerframebox}{\begin{center}\fbox{\parbox{0.92\textwidth}{\BODY}}\end{center}}

\title{Discrete and Computational Geometry \\ Assignment 11}
\author{
  \AA{AAAAAAAAAA AAAAAAA}{6} \\
  \href{mailto:\AA{AAAAAAAAAAAAAAAAAAAA}{7}}{\AA{AAAAAAAAAAAAAAAAAAAA}{7}}
  \and
  Ayush Mishra \\
  \href{mailto:s28amish@uni-bonn.de}{s28amish@uni-bonn.de}
}

\usepackage{graphicx}

\usepackage{titlesec}
\titleformat{\section}
  {\normalfont\Large\bfseries}{Problem \thesection : }
  {0em}{\mdseries}

\begin{document}
  \maketitle
  \begin{center}
    { \bfseries Deadline: 24 Jan 2025, 23:55 }
  \end{center}

  \section{}
  \begin{centerframebox}
    Consider the ground set $\R^2$. Show that
    \begin{enumerate}[label=(\alph*)]
      \item the set system of all axis-parallel squares has VC-dimension at most 3.
      \item the set system of all squares has VC-dimension at least 4.
    \end{enumerate}
  \end{centerframebox}

  \subsection{(a)}
  We need to show that it's impossible to cover all subsets of an arbitrary set of 4 points with axis-aligned squares.
  Sort the points by their $x$ and $y$ coordinates. Mark points $p_{\max x}$, $p_{\min x}$, $p_{\max y}$, $p_{\min y}$.
  The only way for all triplets of points to be covered,
  is for the minimums and maximums of both those lists to be achieved at 4 distinct points.
  Otherwise there would be a point $p_0$, for which no maximum or minimum is achieved.
  A square around the other 3 points would therefore be the bounding box of the whole set, and also include the point $p_0$.

  Now consider how we might cover the pair $\{p_{\max x},\, p_{\min x}\}$.
  The side length of the square covering it must be $\delta_x = \max x - \min x$.
  The square covering $\{p_{\max y},\, p_{\min y}\}$ must have size $\delta_y$ as well.
  One of those deltas must be larger than the other.
  Assume without loss of generality, that $\delta_y \geq \delta_x$.

  Remember that the pair $\{p_{\max y},\, p_{\min y}\}$ must have $p_{\max x}$ to the right of it, and $p_{\min x}$ to the left.
  Because $\delta_x \leq \delta_y$, the horizontal distance between $\{p_{\max y},\, p_{\min y}\}$ and $p_{\max x}$ or $p_{\min x}$ can't be greater than the size of the square covering $\{p_{\max y},\, p_{\min y}\}$.
  So that square must intersect either $p_{\max x}$ or $p_{\min x}$, and it's impossible to squeeze it between them.

  And the VC-dimension must be at most 3.

  \subsection{(b)}
  % Consider the set of 4 points positioned at the corners of a square.
  % Now we try to show that all 16 subsets of those points can be covered by a non-axis-aligned square.
  % The cases with 0 points, 1 point and 4 points are all trivial.

  % For pairs of 2 points, when the form one of the edges of the square, they can also be trivially covered by an axis-aligned square.
  % The other option for
  % TODO: this is wrong


\end{document}
